Multi-target detection with application to cryo-electron microscopy

### 2019-02-15 inverse_problems/inverse_problems.tex
in the middle of Section 2 (skipped section 2)

We consider the \emph{multi-target detection} problem of recovering a set of $K$ signals that appear 
multiple times at unknown locations in a noisy measurement.
Let ${x_1,\ldots,x_K\in\RL}$ be the sought signals and let $y\in\RN$ be the observed data, where we assume $N$ is  far larger than $L$. 
Let  $s[i]$ count the number of signals whose first element is positioned in $y[i]$. Each of those $s[i]$ is chosen according to some (possibly unknown) distribution $\pi$ \TODO{Add notation for the distribution} over $\{1,\ldots,K\}$. 
If signal occurrences overlap, they interfere additively. %We think of those events as being rare. 
With additive white Gaussian noise, the measurement model can be written as 
\begin{equation} 
y  =  \sum_{k=1}^K s_k \ast x_k + \varepsilon, \qquad  \varepsilon   \sim \mathcal{N}(0,\sigma^2 I_N),
\label{eq:model}
\end{equation}
where $\ast$ denotes linear convolution, and $s_k[i]$ indicates the number of starting positions of $x_k$ in $y[i]$ so that $s =  s_1+\cdots+s_K$. Explicitly, with zero-based indexing, % and zero-padding out of bounds,
\begin{align*}
	y[i] & = \sum_{k=1}^{K} \sum_{j = 0}^{L-1} s_k[i-j] x_k[j] + \varepsilon[i].
\end{align*}
The goal is  to estimate $x_1,\ldots,x_K$ from $y$. % while $s$ is the \emph{nuisance variable} of the problem.
%We name this problem \emph{multi-target detection}.
This idealized setup appears in several scientific applications, including structural biology~\cite{bendory2018toward} (as we detail below), spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive}, and system identification~\cite{ljung1998system}.

@@

As a first setup, we allow any generative model for $s$ which meets the following separation requirement: $s$ is binary, and
\begin{equation}
\textrm{If } s[i] = 1 \textrm{ and } s[j] = 1 \textrm{ for } i \neq j, \textrm{ then } |i - j| \geq 2L-1.
\label{eq:spacing}
\end{equation}
In words: the starting positions of any two occurrences  must be separated by at least $2L-1$ positions, so that their end points are necessarily separated by at least $L-1$ signal-free entries in the data.
%Under the separation condition~\eqref{eq:spacing}, we refer to this model as the \emph{homogeneous well-separated model}. 
%In the next section we present a model that alleviates this condition.
%As will be shown next, in a low SNR environment, estimating $s$ is impossible, whereas in some cases estimating $x$ is tractable.
This property ensures that correlating $y$ with versions of itself shifted by at most $L-1$ entries does not involve correlating distinct signal occurrences. Once $s$ is determined, for each position $i$ such that $s[i] = 1$, one of the signals $x_k$ is selected independently at random, and accordingly we set $s_k[i] = 1$. As a result, the only properties of $s_1, \ldots, s_K$ that affect the autocorrelations of $y$ are the total number of occurrences of the distinct signals: their individual and relative locations do not intervene.

@@

In what follows, we consider autocorrelations of both the observation $y$ and of the signal occurrences in $y$. As per our discussion of extensions, the signal occurrences may be sampled from a discrete set $\{x_1, \ldots, x_K\}$, or from a more general distribution. Accordingly, we define autocorrelations for a random signal $z$ of length $m$ \TODO{change $m$ to something upper case, maybe $U$}. For our purposes, this will be applied both to signal occurrences (of length $L$) and to $y$ (of length $N$).

For a random signal $z\in \R^m$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \frac{1}{m} \sum_{i=-\infty}^{\infty} \E\left\{z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}]\right\},
\label{eq:ac_general}
\end{equation}
where the expectation is taken with respect to the distribution of $z$. Indexing of $z$ out of the range $0, \ldots, m-1$ is zero-padded.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align} 
a_z^1 & = \frac{1}{m} \sum_{i=0}^{m-1} \E\left\{
z[i]\right\}, \nonumber\\
a_z^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} \E\left\{z[i]z[i+\ell]\right\},\\
a_z^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} \E\left\{z[i]z[i+\ell_1]z[i+\ell_2]\right\}.  \nonumber \label{eq:ac_special}
\end{align}
Since  autocorrelations depend only on the differences between indices, the obey the following symmetries: $$a_z^2[\ell] = a_z^2[-\ell],$$ and
$$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$$

Under the separation condition, the relation between autocorrelations of the micrograph and those of $X$ is particularly simple, as we now show. It is useful to introduce some notation: let $M$ denote the number of occurrences of $X$ in $y$ (that is, the number of 1's in $s$), and let
\begin{equation}
\gamma  = \frac{M L}{N}
\label{eq:gamma}
\end{equation}
denote the density of $X$ in $y$ (that is, the fraction of entries of $y$ occupied by occurrences of $X$.) 
In this section, we assume the separation condition~\eqref{eq:spacing} that  imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$. In Section~\ref{sec:poisson} we show that the autocorrelations in the Poisson model are equivalent to  those presented in this section.

For shifts in $0, \ldots, L-1$, the autocorrelation functions of $y$ depend on the corresponding autocorrelations of $x$, the noise level $\sigma$ and the support signal $s$. Importantly, under the separation condition~\eqref{eq:spacing}, the dependency on $s$ is only through the density $\gamma$.
We consider the asymptotic regime where $\gamma$ remains constant; that is, as $N$ goes to infinity, $M$ also goes to infinity at the same rate (in other words, as we see an increasingly large observation, it contains increasingly many signal occurrences, with constant signal density). In that regime, the law of large numbers can be used to show the following statement:
\begin{equation} \label{eq:mean_micrograph}
\lim_{N\to\infty} a_y^1  \aseq \gamma a_{X}^1,
\end{equation}
\TODO{ should is be $\E\left\{a_{X}^1\right\}$?} where equality holds almost surely (a.s.), meaning it holds with probability one.
The randomness is over the Gaussian noise $\varepsilon$; $s$ may be deterministic.
Thus, given enough data, if $\gamma$ is known, we can estimate $a_{X}^1$ from $y$. We show in Section~\ref{sec:theory_homogeneous} how to estimate $\gamma$ as well in the homogeneous model.

@@

In particular, for $X$ sampled from $\{ x_1, \ldots, x_K \}$ with probabilities $(\pi_1, \ldots, \pi_K)$, the autocorrelations are given in explicit form as:
\begin{align} 
a_X^1 & = \frac{1}{m} \sum_{i=0}^{m-1} \E\left\{
X[i]\right\}, \nonumber\\
a_X^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} \E\left\{X[i]X[i+\ell]\right\},\\
a_X^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} \E\left\{X[i]X[i+\ell_1]X[i+\ell_2]\right\}.  \nonumber
\end{align}
\TODO{ABOVE: replace expectations with sums.}

@@

@@

\appendix

\section{Derivation of the identities in Section~\ref{sec:AC_analysis}} \label{sec:autocorrelation_computation}

We consider the asymptotic regime $N,M\to\infty$ and assume that $M=\Omega(N)$, so that
\begin{equation*}
\gamma := \frac{ML}{N}>0.
\end{equation*}
%
Any vectors with indices out of range are given value 0.
We focus here on the homogeneous model~\eqref{eq:model}. The extension to random signal $X$ is straight-forward by taking the expectation with respect to the distribution of $X$.

We start by considering the mean of the data:
\begin{equation*}
a_y^1 = \frac{1}{N}\sum_{i=0}^{N-1} y[i] =
\frac{1}{N/L}\sum_{j=0}^{M-1}\frac{1}{L}\sum_{i=0}^{L-1}x[i] +    
\underbrace{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]}_{\text{noise term}}
\xrightarrow{a.s.}\gamma a_x^1,
\end{equation*}
%
where the noise term converges to zero almost surely (a.s.)\ by the strong law of large numbers.

We proceed with the (second-order) autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
%
\begin{align*}
%
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]
\nonumber \\
& = \underbrace{\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}}
\\ & + \underbrace{\frac{1}{N} \sum_{j=1}^{M} \sum_{i=0}^{L-1} x[i] (\varepsilon[s_j + i + \ell] + \varepsilon[s_j + i - \ell])}_{\text{cross-term}}. 
%
\end{align*}
The cross-term is linear in the noise, and is easily shown to vanish almost surely in the limit $N\to\infty$, by the strong law of large numbers. We break the signal term into $M$ different sums, each containing one copy of the signal. This gives:
%
\begin{align} \label{eq:2nd_moment_signal_term}
%
\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] & = \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\nonumber\\&\xrightarrow{N\to\infty}\gamma a_x^2[\ell].
%
\end{align}
%
We next analyze the pure noise term. When $\ell\neq 0$, we can break the noise term into a sum of independent terms:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -2} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
%
\end{equation*}
%
Each sum $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -2} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, hence converges to zero almost surely as $N\to\infty$. If $\ell=0$, then we have:
%
\begin{equation*}
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
\end{equation*}

We now analyze the third-order autocorrelation. Let us fix $\ell_1\geq\ell_2\ge0$. We have:
%
%\begin{align*}
%%
%&a_y^3[\ell_1,\ell_2] 
%= \frac{1}{N}\sum_{i=0}^{N-1-\ell_1} y[i]y[i+\ell_1]y[i+\ell_2]
%\nonumber \\
%%
%=& \underbrace{ \frac{ML}{N}\frac{1}{M}\sum_{j=1}^M 
%	\frac{1}{L}\sum_{i=0}^{L-1-\ell_1}x[i]x[i+\ell_1]x[i+\ell_2]   }_{(1)}
%\\& + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} \ep[i]\ep[i+\ell_1]\ep[i+\ell_2]}_{(2)}
%\nonumber \\
%&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
%	\sum_{i=0}^{L-1} x[i]\ep[s_j + i+\ell_1]\ep[s_j+ i+\ell_2]}_{(3)}
%\\& + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
%	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]}_{(4)}
%\nonumber \\
%&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
%	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_2]\ep[s_j+i+\ell_1-\ell_2]x[i]}_{(5)}
%\\ & + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
%	\sum_{i=0}^{L-\ell_1+\ell_2-1} \ep[s_j+i-\ell_2]x[i+\ell_1-\ell_2]x[i]}_{(6)}
%\nonumber \\
%&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
%	\sum_{i=0}^{L-\ell_2-1} x[i]\ep[s_j + i+\ell_1]x[i+\ell_2]}_{(7)}
%\\ & + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
%	\sum_{i=0}^{L-\ell_1-1} x[i]x[i+\ell_1]\ep[s_j+ i+\ell_2]}_{(8)}.
%%
%\end{align*}
%
\begin{align*}
%
a_y^3[\ell_1,\ell_2] 
=& \frac{1}{N}\sum_{i=0}^{N-1-\ell_1} y[i]y[i+\ell_1]y[i+\ell_2]
\\
%
=& \underbrace{ \frac{ML}{N}\frac{1}{M}\sum_{j=1}^M 
	\frac{1}{L}\sum_{i=0}^{L-1-\ell_1}x[i]x[i+\ell_1]x[i+\ell_2]   }_{(1)}
+ \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} \ep[i]\ep[i+\ell_1]\ep[i+\ell_2]}_{(2)}
\ \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} x[i]\ep[s_j + i+\ell_1]\ep[s_j+ i+\ell_2]}_{(3)}
+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]}_{(4)}
\\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_2]\ep[s_j+i+\ell_1-\ell_2]x[i]}_{(5)}
+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_1+\ell_2-1} \ep[s_j+i-\ell_2]x[i+\ell_1-\ell_2]x[i]}_{(6)}
\\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_2-1} x[i]\ep[s_j + i+\ell_1]x[i+\ell_2]}_{(7)}
+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_1-1} x[i]x[i+\ell_1]\ep[s_j+ i+\ell_2]}_{(8)}.
%
\end{align*}
Terms (6), (7) and (8) are linear in $\ep$, and can easily be shown to converge to 0 almost surely by the law of large numbers, by similar arguments as used previously. Term (1) converges to $\gamma a_x^3[\ell_1,\ell_2]$ almost surely, for the same reasons as~\eqref{eq:2nd_moment_signal_term}. To deal with terms (2)--(5), we must distinguish between different values of $\ell_1$ and $\ell_2$.

{\bf Case 1:} $0 < \ell_2 < \ell_1$. Here, all summands with elements of $\ep$ involve products of distinct entries, which have expected value 0. Consequently, the usual argument shows that terms (2)--(5) all converge to 0 almost surely as $N \to \infty$.

{\bf Case 2:} $0=\ell_2 < \ell_1$. Term (2) is an average of products of the form $\ep[i]^2\ep[i+\ell_1]$, which have mean zero; consequently, term (2) converges to 0 almost surely. The same argument as for Case 1 shows that (3) and (5) also converge to 0. For term (4), we write:
%
\begin{align*}
%
&\frac{1}{N}\sum_{j=1}^{M} 
\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]
\nonumber \\
&= \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-1}x[i] \frac{1}{M}\sum_{j=1}^{M} \ep[s_j+i-\ell_1]^2
\\& \xrightarrow{a.s.} \gamma \frac{1}{L} \sum_{i=0}^{L-1}x[i] \sigma^2 = \gamma a_x^1 \sigma^2. \nonumber
%
\end{align*}

{\bf Case 3:} $0<\ell_2 = \ell_1$. An argument nearly identical to that for Case 2 shows that terms (2), (4) and (5) converge to 0, while term (3) converges to $\gamma a_x^1 \sigma^2$.

{\bf Case 4:} $0=\ell_2 = \ell_1$. The same argument as for term (4) in Case 2 shows that terms (3), (4) and (5) all converge to $\gamma a_x^1 \sigma^2$. Term (2) is an average of $\ep[i]^3$, which is mean zero; consequently, it converges to 0.
This completes the proof.

\section{Proof of Proposition~\ref{prop:gamma}} \label{sec:proof_prop_gamma}

\section{Proof of Proposition~\ref{prop:gamma_sigma}} \label{sec:proof_prop_gamma_sigma}

\section{Proof of Proposition~\ref{prop:poisson}} \label{sec:proof_prop_poisson}

\subsection{First moment}

\subsection{Second moment}

\subsection{Third moment}

>>>>

We consider the \emph{multi-target detection} problem of recovering a set of $K$ signals that appear 
multiple times at unknown locations in a noisy measurement.
Let ${x_1,\ldots,x_K\in\RL}$ be the sought signals and let $y\in\RN$ be the observed data, where we assume $N$ is  far larger than $L$. 
Let  $s[i]$ count the number of signals whose first element is positioned in $y[i]$. Each of those $s[i]$ is chosen according to some (possibly unknown) distribution over $\{1,\ldots,K\}$. 
If signal occurrences overlap, they interfere additively. %We think of those events as being rare. 
With additive white Gaussian noise, the measurement model can be written as 
\begin{equation} 
y  =  \sum_{k=1}^K s_k \ast x_k + \varepsilon, \qquad  \varepsilon   \sim \mathcal{N}(0,\sigma^2 I_N),
\label{eq:model}
\end{equation}
where $\ast$ denotes linear convolution, and $s_k[i]$ indicates the number of starting positions of $x_k$ in $y[i]$ so that $s =  s_1+\cdots+s_K$. Explicitly, with zero-based indexing, % and zero-padding out of bounds,
\begin{align*}
	y[i] & = \sum_{k=1}^{K} \sum_{j = 0}^{L-1} s_k[i-j] x_k[j] + \varepsilon[i].
\end{align*}
The goal is  to estimate $x_1,\ldots,x_K$ from $y$. % while $s$ is the \emph{nuisance variable} of the problem.
%We name this problem \emph{multi-target detection}.
This idealized setup appears in several scientific applications, including structural biology~\cite{bendory2018toward} (as we detail below), spike sorting~\cite{lewicki1998review}, passive radar~\cite{gogineni2017passive}, and system identification~\cite{ljung1998system}.

@@

As a first setup, we allow any generative model for $s$ which meets the following separation requirement: $s$ is binary, and
\begin{equation}
\textrm{If } s[i] = 1 \textrm{ and } s[j] = 1 \textrm{ for } i \neq j, \textrm{ then } |i - j| \geq 2L-1.
\label{eq:spacing}
\end{equation}
In words: the starting positions of any two occurrences  must be separated by at least $2L-1$ positions, so that their end points are necessarily separated by at least $L-1$ signal-free entries in the data.
%Under the separation condition~\eqref{eq:spacing}, we refer to this model as the \emph{homogeneous well-separated model}. 
%In the next section we present a model that alleviates this condition.
%As will be shown next, in a low SNR environment, estimating $s$ is impossible, whereas in some cases estimating $x$ is tractable.
Furthermore, we require that the last signal occurrence in $y$ is also followed by at least $L-1$ signal-free (but still noisy) entries.
% This last sentence is required for the proof of the identity of the third-order autocorrelation in Section 3.
This property ensures that correlating $y$ with versions of itself shifted by at most $L-1$ entries does not involve correlating distinct signal occurrences. Once $s$ is determined, for each position $i$ such that $s[i] = 1$, one of the signals $x_k$ is selected independently at random, and accordingly we set $s_k[i] = 1$. As a result, the only properties of $s_1, \ldots, s_K$ that affect the autocorrelations of $y$ are the total number of occurrences of the distinct signals: their individual and relative locations do not intervene.

@@

In what follows, we consider autocorrelations of both the observation $y$ and of the signal occurrences in $y$. As per our discussion of extensions, the signal occurrences may be sampled from a discrete set $\{x_1, \ldots, x_K\}$, or from a more general distribution. Accordingly, we define autocorrelations for a random signal $z$ of length $M$ . For our purposes, this will be applied both to signal occurrences (of length $L$) and to $y$ (of length $N$).

For a random signal $z\in \R^M$, the autocorrelation of order $q = 1, 2, \ldots$ is given for any integer shifts $\ell_1, \ldots, \ell_{q-1}$ by
\begin{equation}
a_z^q[\ell_1,\ldots,\ell_{q-1}]   = \E_z\left\{\frac{1}{M} \sum_{i=-\infty}^{\infty} z[i]z[i+\ell_1]\cdots z[i+\ell_{q-1}]\right\},
\label{eq:ac_general}
\end{equation}
where the expectation is taken with respect to the distribution of $z$. Indexing of $z$ out of the range $0, \ldots, M-1$ is zero-padded.
Explicitly, the first-, second- and third-order autocorrelations are given by: 
\begin{align} 
a_z^1 & = \E_z\left\{\frac{1}{M} \sum_{i=0}^{M-1} 
z[i]\right\}, \nonumber\\
a_z^2[\ell] & = \E_z\left\{\frac{1}{M} \sum_{i = \max\{0, -\ell\}}^{M-1 + \min\{0, -\ell\}} z[i]z[i+\ell]\right\},\\
a_z^3[\ell_1,\ell_2] & = \E_z\left\{\frac{1}{M} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{M-1 + \min\{0, -\ell_1, -\ell_2\}} z[i]z[i+\ell_1]z[i+\ell_2]\right\}.  \nonumber \label{eq:ac_special}
\end{align}
Since  autocorrelations depend only on the differences between indices, the obey the following symmetries: $$a_z^2[\ell] = a_z^2[-\ell],$$ and
$$a_z^3[\ell_1,\ell_2] = a_z^3[\ell_2,\ell_1]=a_z^3[-\ell_1,\ell_2-\ell_1].
$$

Let $x$ be the random variable corresponding to signal occurrences in $y$. In particular, for the model~\eqref{eq:model}, $x=x_k$ with probability $\pi_k$.
Under the separation condition~\eqref{eq:spacing}, the relation between autocorrelations of the observation $y$ and those of $x$ is particularly simple, as we now show. It is useful to introduce some notation: let $\vert s\vert = \sum_i s[i]$ denote the number of signal occurrences in $y$, and let
\begin{equation}
\gamma  = \frac{|s| L}{N}.
\label{eq:gamma}
\end{equation}
In the well-separated model, $\gamma$ is the fraction of entries of $y$ occupied by signal occurrences. Henceforth, we call it   
 the signal density. The separation condition~\eqref{eq:spacing}  imposes $\gamma\leq\frac{L}{2L-1}\approx 1/2$. 
 In Section~\ref{sec:poisson} we show that the autocorrelations under the Poisson model are equivalent to those presented in this section.

For shifts in $0, \ldots, L-1$, the autocorrelations of $y$ depend on the corresponding autocorrelations of $x$, the noise level $\sigma$ and the support signal $s$. 
%Importantly, under the separation condition~\eqref{eq:spacing}, the dependency on $s$ is only through the density $\gamma$.
We consider the asymptotic regime where $\gamma$ remains constant; that is, as $N$ goes to infinity, $|s|$ also goes to infinity at the same rate. %(in other words, as we see an increasingly large observation, it contains increasingly many signal occurrences, with constant signal density). 
In that regime, the law of large numbers implies
\begin{equation} \label{eq:mean_micrograph}
\lim_{N\to\infty} a_y^1  \aseq \gamma a_{x}^1,
\end{equation}
where equality holds almost surely (a.s.), meaning it holds with probability one.
The randomness is over $x$ and the Gaussian noise $\varepsilon$; $s$ may be deterministic.
Thus, given enough data, if $\gamma$ is known, we can estimate $a_{x}^1$ from $y$. We show in Section~\ref{sec:theory_homogeneous} how to estimate $\gamma$ as well in a particular case.

@@

In particular, for $x$ sampled from $\{ x_1, \ldots, x_K \}$ with probabilities $(\pi_1, \ldots, \pi_K)$, the autocorrelations are given in explicit form as:
\begin{equation}
a_x^q  =  \sum_{i=1}^{K} \pi_ka_{x_i}^1.
\end{equation}
%\begin{align} 
%a_X^1 & = \frac{1}{K} \sum_{i=1}^{K} \pi_ka_{x_i}, \nonumber\\
%a_X^2[\ell] & = \frac{1}{m} \sum_{i = \max\{0, -\ell\}}^{m-1 + \min\{0, -\ell\}} \E\left\{X[i]X[i+\ell]\right\},\\
%a_X^3[\ell_1,\ell_2] & = \frac{1}{m} \sum_{i = \max\{0, -\ell_1, -\ell_2\}}^{m-1 + \min\{0, -\ell_1, -\ell_2\}} \E\left\{X[i]X[i+\ell_1]X[i+\ell_2]\right\}.  \nonumber
%\end{align}
%\TODO{ABOVE: replace expectations with sums.}

@@

\section{Proof}

\subsection{Derivation of the identities in Section~\ref{sec:AC_analysis}} \label{sec:autocorrelation_computation}
Let us define
\begin{equation*}
\gamma = \frac{|s|L}{N}>0.
\end{equation*}
Then the first order autocorrelation of the observation is given by 
\begin{equation*}
a_y^1 = \E_y\left\{\frac{1}{N}\sum_{i=0}^{N-1} y[i]\right\} =
\frac{|s|}{N}\E_x\left\{\sum_{i=0}^{L-1}x[i]\right\} +    
\E_{\varepsilon}\left\{\frac{1}{N}\sum_{i=0}^{N-1}\varepsilon[i]\right\}
=\gamma a_x^1,
\end{equation*}
where $\E_z$ means expectation with respect to the distribution of $z$.
%the noise term converges to zero almost surely (a.s.)\ by the strong law of large numbers.

We proceed with the second-order autocorrelation for  $\ell\in\{0,\ldots,L-1\}$. We can compute:
%
\begin{align*}
%
a_y^2[\ell] & = \E_y\left\{\frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]\right\}
\nonumber \\
& = \frac{|s|}{N}\E_x\left\{\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\right\}  + \E_{\varepsilon}\left\{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell] \right\},
%
\end{align*}
where all cross terms vanish in expectation owing to independence of $x$ of $\varepsilon$, and $\E{\varepsilon}=0$.  
Hence, 
\begin{align} \label{eq:2nd_moment_signal_term}
%
a_y^2[\ell] & = \gamma a_x^2[\ell] + \frac{N-1-\ell}{N}\sigma^2\delta[\ell],
\end{align}
where $\delta[0]=1$ and zero otherwise.

We now analyze the third-order autocorrelation. 
Let $x_{(j)}$ denote the $j$th
realization of $x$, with starting position $s_{(j)}$, for $j=1,\ldots,|s|$.
Let us fix $0\leq \ell_2\leq\ell_1\leq L-1$. We have:
\begin{align*}
%
a_y^3[\ell_1,\ell_2] 
=& \E_y\left\{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} y[i]y[i+\ell_1]y[i+\ell_2]\right\}
\\
%
\stackrel{(1)}=& \E_{x_{(1)},\ldots,x_{(|s|)}}\left\{ \frac{|s|L}{N}\frac{1}{|s|}\sum_{j=1}^{|s|} 
	\frac{1}{L}\sum_{i=0}^{L-1-\ell_1}x_{(j)}[i]x_{(j)}[i+\ell_1]x_{(j)}[i+\ell_2]  \right\} \\
\stackrel{(2)}+& \E_\varepsilon\left\{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} \ep[i]\ep[i+\ell_1]\ep[i+\ell_2]\right\}
\ \\
\stackrel{(3)}+& \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
	\sum_{i=0}^{L-1} x_{(j)}[i]\ep[s_{(j)} + i+\ell_1]\ep[s_{(j)}+ i+\ell_2]\right\} \\
\stackrel{(4)}+& \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
	\sum_{i=0}^{L-1} \ep[s_{(j)}+i-\ell_1]x_{(j)}[i]\ep[s_{(j)}+ i+\ell_2-\ell_1]\right\}
\\
\stackrel{(5)}+& \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
	\sum_{i=0}^{L-1} \ep[s_{(j)}+i-\ell_2]\ep[s_{(j)}+i+\ell_1-\ell_2]x_{(j)}[i]\right\}
\\ \stackrel{(6)}+ & \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
	\sum_{i=0}^{L-\ell_1+\ell_2-1} \ep[s_{(j)}+i-\ell_2]x_{(j)}[i+\ell_1-\ell_2]x_{(j)}[i]\right\}
\\
\stackrel{(7)}+ & \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
	\sum_{i=0}^{L-\ell_2-1} x_{(j)}[i]\ep[s_{(j)} + i+\ell_1]x_{(j)}[i+\ell_2]\right\}
\\ \stackrel{(8)}+ & \E_{x_{(1)},\ldots,x_{(|s|)},\varepsilon}\left\{\frac{1}{N}\sum_{j=1}^{|s|} 
	\sum_{i=0}^{L-\ell_1-1} x_{(j)}[i]x_{(j)}[i+\ell_1]\ep[s_{(j)}+ i+\ell_2]\right\}.
%
\end{align*}
Terms (6), (7) and (8) are linear in $\ep$ and  thus zero.
Term (1) is equal  to $\gamma a_x^3[\ell_1,\ell_2]$  for the same reasons as~\eqref{eq:2nd_moment_signal_term}. 
Term (2) is always zero by properties of Gaussian variables. 
To deal with terms (3)--(5), we must distinguish between different values of $\ell_1$ and $\ell_2$.

Term (3) is always zero, unless $\ell_1=\ell_2$, so that it is equal to 
$\gamma\sigma^2 a_x^1 \delta[\ell_1-\ell_2]$.
The same goes for Terms (4) and (5), which are equal, respectively, to
$\gamma\sigma^2 a_x^1 \delta[\ell_2]$ 
and $\gamma\sigma^2 a_x^1 \delta[\ell_1]$.

\subsection{Proof of Proposition~\ref{prop:gamma}} \label{sec:proof_prop_gamma}

\subsection{Proof of Proposition~\ref{prop:gamma_sigma}} \label{sec:proof_prop_gamma_sigma}

\subsection{Proof of Proposition~\ref{prop:poisson}} \label{sec:proof_prop_poisson}

\subsubsection{First moment}

\subsubsection{Second moment}

\subsubsection{Third moment}

@@
